### 2025-03-26 ms/ms_old.tex
update data availability section

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, enhanced, frame hidden, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Mapping terrestrial carbon stocks is crucial to successfully
implementing climate change mitigation programs
(\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}). The accuracy of
carbon stock estimates depends on the availability of reliable
allometric models to deduce the aboveground biomass (AGB) of trees from
census data (\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}).
Selecting an appropriate allometric model is, therefore, a critical step
in estimating forest carbon biomass (e.g.,
\protect\hyperlink{ref-Chave2014}{Chave et al. 2014};
\protect\hyperlink{ref-Molto2014}{Molto et al. 2014}). Tree allometry
describes the correlations between different biometric measurements of
trees, such as diameter at breast height (DBH), tree height, and tree
crown size (e.g., \protect\hyperlink{ref-Shinozaki1964a}{Shinozaki et
al. 1964}; \protect\hyperlink{ref-Niklas1994}{Niklas 1994};
\protect\hyperlink{ref-West1999}{West et al. 1999}). These relationships
are widely utilized to estimate tree architecture from DBH and to
estimate forest biomass and production, serving as fundamental
components of dynamic vegetation models (e.g.,
\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
\protect\hyperlink{ref-Falster2017}{Falster et al. 2017};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}). Despite significant
progress, developing precise allometric models across regions
characterized by diverse tree species composition and varying forest
structures remains a persistent challenge. This complexity highlights
the need for improved global-scale allometric models that capture the
diversity of tree forms and ecological contexts.

Identifying the most appropriate mathematical representation of
allometric relationships is a subject of ongoing debate. While numerous
papers have used various functional forms for allometric equations
(e.g., \protect\hyperlink{ref-Thomas1996}{Thomas 1996};
\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}), power-law functions
remain the most common choice for describing the allometric scaling of
tree size variables, including height and crown size with DBH (e.g.,
\protect\hyperlink{ref-Sileshi2023}{Sileshi et al. 2023};
\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}). The
relationships among tree size variables often involve complex
interactions influenced by various factors, such as species diversity,
environmental conditions, and individual variability
(\protect\hyperlink{ref-Laurans2024}{Laurans et al. 2024}; e.g.,
\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}).
Consequently, power-law models often fail to accurately capture the
allometric relationships across the full range of tree sizes, leading to
underestimation of dimensions for smaller trees and underestimation for
larger trees (e.g., \protect\hyperlink{ref-Fayolle2016}{Fayolle et al.
2016}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). This
highlights the need to adopt alternative functional forms---such as
saturating relationships---that can more accurately reflect the
complexity of allometric scaling in vegetation models (e.g.,
\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}).

Earth observation (EO) technologies, including remote sensing and Light
Detection and Ranging (LiDAR), have revolutionized monitoring forests
(e.g., \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}) by
enabling the identification and measurement of individual tree heights
and crown dimensions using airborne imagery (e.g.,
\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). Because
existing allometric models still depend on stem diameter as a crucial
input for biomass estimation, it also presents the challenge of
determining the most effective way to utilize EO technologies for AGB
estimation (\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). While
ground-based LiDAR can measure DBH directly and is much more efficient
than traditional ground surveys
(\protect\hyperlink{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is still
less common and convenient than aerial LiDAR. Alternative approaches
focusing on tree height and crown dimensions as model predictors are
also necessary for remote sensing applications
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). One potential
solution is to estimate DBH from tree height and crown size and then use
established biomass equations, which could be an effective strategy for
integrating airborne imagery into carbon monitoring programs
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).

Accurately representing tree allometry relationships also requires
considering underlying ecological constraints, particularly the
interplay of plant functional traits and their trade-offs influences
individual tree architecture and the composition of entire forest
communities (e.g., \protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). Under
closed-canopy conditions, strong light competition makes the balance
between vertical and horizontal space use crucial for species
coexistence (e.g., \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka
2009}; \protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Among
functional traits, wood density is closely linked to both plant life
history strategies
(\protect\hyperlink{ref-Fajardo2022}{\textbf{Fajardo2022?}}) and light
demand (e.g., \protect\hyperlink{ref-King2006}{King et al. 2006};
\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}). Several
studies suggest that wood density explains a trade-off between rapid
stem expansion and broader crowns, shaping mechanical constraints on
species with diverse architecture and light capture strategies
(\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). However, these
types of studies have examined this relationship post hoc, relating
species-specific allometric parameters to wood density and risking Type
I or Type II errors if the hierarchical data structure is overlooked
(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}). One
exception is Martínez Cano et al.
(\protect\hyperlink{ref-MartinezCano2019}{2019}), who incorporated wood
density into a hierarchical model of tropical trees and concluded it
offered minimal explanatory power for interspecific variation.
Therefore, the role of wood density in global allometric scaling remains
insufficiently explored.

In this study, we present a robust Bayesian hierarchical approach to
delineate allometric relationships involving DBH, tree height (H), crown
radius (CR), and their interspecific variations. Using large datasets
from the Tallo and TRY databases
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020};
\protect\hyperlink{ref-Jucker2022}{Jucker et al. 2022}), which span a
diverse array of tree forms across global forest types and climatic
conditions, we overcome the limitations of previous studies that often
relied on site-specific or regional observations with limited samples
and species. By integrating species-level wood density into a multilevel
model, our methodology avoids the post-hoc correlations used in previous
studies, offering a more universal and scalable model, effectively
capturing variability from the species to the community level.
Specifically, we ask the following questions: (1) What is the best
function to estimate tree architecture from DBH and \emph{vice versa}?
(2) How does wood density shape interspecific variation in these
allometric scaling relationships? (3) How does the choice of allometric
relationships affect subsequent AGB estimation?

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

Our analysis combined tree allometry data from Tallo, which includes
standardized measurements of DBH, height, crown radius for 5,163 species
across various biomes (\protect\hyperlink{ref-Jucker2022}{Jucker et al.
2022}, https://doi.org/10.5281/zenodo.6637599) along with wood density
data for 10,995 species from TRY version 6
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020},
https://www.try-db.org/TryWeb/Home.php) (see Supplementary Table S1 for
further details). We thoroughly processed the Tallo dataset by removing
998 outliers following Jucker et al.
(\protect\hyperlink{ref-Jucker2022}{2022}). To ensure consistency, we
selected only species with standardized wood density values in g
cm\textsuperscript{-3} from TRY and removed any records with missing
data. Subsequently, we created an overlapping dataset between Tallo and
TRY, resulting in a combined dataset encompassing allometric and wood
density information for 355,881 individuals across 2,961 species.

To reduce computational complexity and ensure the robustness of the
models, we further filtered species with fewer than 20 individuals and
randomly sampled up to 100 individuals per species. This step was
necessary because the dataset is unbalanced; for example, \emph{Quercus
ilex} alone accounts for over 20,000 individuals, representing
approximately 10\% of the dataset. However, \emph{Quercus ilex} is not
necessarily the most abundant species globally. Its prevalence in this
dataset is a result of sampling and does not reflect its true global
abundance. By limiting the number of individuals per species, we
minimize bias due to data imbalance and prevent the overrepresentation
of any single abundant species
(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}), thereby
ensuring that the model remains broadly applicable. Since not all trees
have a complete set of information, we created specific sub-datasets to
select the best predictive model for each allometric relationship for
two clades (i.e., angiosperms and gymnosperms; see below and
Supplementary Table S2 for more details).

\hypertarget{tree-allometry-models}{%
\subsection{Tree allometry models}\label{tree-allometry-models}}

@@

To compare the performance of different models for the same response
variable, we used Pareto-smoothed importance sampling leave-one-out
cross-validation (PSIS-LOO; Vehtari et al.
(\protect\hyperlink{ref-Vehtari2017}{2017})). PSIS-LOO is an accurate
and reliable approximation to standard leave-one-out cross-validation
(LOO), which is a robust method for comparing models with different
numbers of parameters (\protect\hyperlink{ref-Vehtari2017}{Vehtari et
al. 2017}). We utilized PSIS-LOO to calculate the LOO Information
Criterion (LOOIC) for each model. A model with lower LOOIC is considered
to have better predictive accuracy
(\protect\hyperlink{ref-Vehtari2017}{Vehtari et al. 2017}).

Once the best-performing models were identified for each allometric
relationship, we then tested whether adding a species-level predictor
(wood density) would further improve that model's predictive
performance. Species-specific parameters (\(\beta_j\)) were modeled as a
function of wood density (\(u_j\)):
\begin{equation}\protect\hypertarget{eq-wood}{}{
\beta_{j} \sim N(\gamma_{0} + \gamma_{1} u_j, \sigma_{\beta}),
}\label{eq-wood}\end{equation}

@@

Posterior distributions of all parameters were estimated using the
No-U-Turn Sampler (NUTS), an adaptive variant of the Hamiltonian Monte
Carlo (HMC), algorithm implemented in Stan
(\protect\hyperlink{ref-Carpenter2017}{Carpenter et al. 2017}) using
weakly-informative priors (\protect\hyperlink{ref-Gelman2008}{Gelman et
al. 2008}). The HMC algorithm uses gradient information to propose new
states in the Markov chain, leading to a more efficient exploration of
the target distribution than traditional Markov chain Monte Carlo (MCMC)
methods that rely on random proposals
(\protect\hyperlink{ref-Carpenter2017}{Carpenter et al. 2017}). This
efficiency allows us to achieve convergence with fewer iterations than
traditional MCMC methods. Inferences were based on 1,000 posterior
samples following 1,000 warmup iterations for three parallel chains.
Convergence was assessed using the Gelman--Rubin statistic with a
convergence threshold of 1.1 (\protect\hyperlink{ref-Gelman2013}{Gelman
et al. 2015}) and ensuring effective sample sizes greater than 300
(\protect\hyperlink{ref-Vehtari2021}{Vehtari et al. 2021}). We also
reported the median and 95\% Bayesian credible intervals (CIs) for each
parameter (Supplementary Tables S3-S8). Estimates were considered
significant when 95\% CIs did not overlap with zero.

Statistical analyses were performed in R version 4.3.2
(\protect\hyperlink{ref-RCoreTeam2023}{R Core Team 2023}), using the R
packages \emph{targets} version 1.7.0 for workflow management
(\protect\hyperlink{ref-Landau2021}{Landau 2021}). Codes are made
available on GitHub at
\url{https://github.com/duyenecology/tree-allometry} (Stan code is
provided in Supplementary Section S2).

\hypertarget{aboveground-biomass-estimates}{%
\subsection{Aboveground biomass
estimates}\label{aboveground-biomass-estimates}}

For angiosperm species, we estimated aboveground biomass (AGB; kg of dry
mass) using the equation from Chave et al.
(\protect\hyperlink{ref-Chave2014}{2014}):

\begin{equation}\protect\hypertarget{eq-agb}{}{
\mathrm{AGB} = 0.0673 \times (\rho D^2H)^{0.976}
}\label{eq-agb}\end{equation}

where \(\rho\) is wood density (g cm\textsuperscript{-3}), \(D\) (cm) is
DBH, and \(H\) (m) is tree height. We either used both observed DBH and
height or substituted one of these variables with an estimate from our
allometric models, allowing us to assess how different allometric
relationships affect AGB calculations. Only trees with a DBH ≥ 5 cm were
used, as those below this threshold hold a small fraction of forest
carbon stocks and were excluded from the calibration of Chave et al.
(\protect\hyperlink{ref-Chave2014}{2014})'s equation. Additionally, the
sub-datasets used to select the optimal predictive model for each
allometric relationship differed as not all trees had a complete
information set (see above and Supplementary Table S2). Consequently,
the final dataset for estimating AGB in angiosperms includes 124,206
trees with sufficient allometric information (i.e., DBH, H, CR, WD)
across 624 species. We did not extend this approach to gymnosperms
because no general AGB equation is available for them, and most existing
equations are species-specific and do not incorporate wood density
(e.g., \protect\hyperlink{ref-Moore2010}{Moore 2010};
\protect\hyperlink{ref-Bazrgar2024}{Bazrgar et al. 2024}).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{tree-height-allometry}{%
\subsection{Tree height allometry}\label{tree-height-allometry}}

The optimal model to predict tree height was a Weibull function for both
angiosperms (ang) and gymnosperms (gym) (Table 1). Across the global
dataset, the best model to predict H from DBH of each division was:
\begin{equation}\protect\hypertarget{eq-h_ang}{}{
\mathrm{H_{ang}} =50.0
\left\{ 1 - \exp\left( - 0.0476
\mathrm{DBH}^{0.629}
\right) \right\}
}\label{eq-h_ang}\end{equation}

\begin{equation}\protect\hypertarget{eq-h_gym}{}{
\mathrm{H_{gym}} = 117.4
\left\{ 1 - \exp\left( - 0.012
\mathrm{DBH}^{0.750}
\right) \right\}.
}\label{eq-h_gym}\end{equation}

@@

{\centering \includegraphics{../figs/h_cr_dbh_non_log.png}

@@

\end{figure}

\hypertarget{crown-radius-allometry}{%
\subsection{Crown Radius allometry}\label{crown-radius-allometry}}

The power-law function (0 \textless{} \emph{b} \textless{} 1) provided
the best predictive performance for modeling the relationship between
crown radius and DBH in angiosperms. Whereas for gymnosperms, the gMM
function offered the best performance for this allometric scaling (Table
1). Across the global dataset, the best model for predicting CR from DBH
of each division was: \begin{equation}\protect\hypertarget{eq-cr_ang}{}{
\mathrm{CR_{ang}} = \mathrm{0.349}
\, \mathrm{DBH}^{\mathrm{0.626}}
}\label{eq-cr_ang}\end{equation}

\begin{equation}\protect\hypertarget{eq-cr_gym}{}{
\mathrm{CR_{gym}} = \frac{\mathrm{769.4}
\, \mathrm{DBH}^{\mathrm{0.631}}}
{\mathrm{2657}
+ \mathrm{DBH}^{\mathrm{0.631}}}.
}\label{eq-cr_gym}\end{equation}

@@

\hypertarget{dbh-allometry}{%
\subsection{DBH allometry}\label{dbh-allometry}}

The \(CR,H\) model, which treated CR and H as separate factors,
exhibited the best predictive performance among all tested models (Table
1). Across the global dataset, the most effective model for predicting
DBH from H and CR for each division was:
\begin{equation}\protect\hypertarget{eq-dbh_ang}{}{
\mathrm{DBH_{ang}} = \mathrm{0.948}
\mathrm{CR}^{\text{0.374}}
\mathrm{H}^{\text{0.590}}
}\label{eq-dbh_ang}\end{equation}

\begin{equation}\protect\hypertarget{eq-dbh_gym}{}{
\mathrm{DBH_{gym}} = \mathrm{0.972}
\mathrm{CR}^{\text{0.410}}
\mathrm{H}^{\text{0.588}}.
}\label{eq-dbh_gym}\end{equation}

@@

\hypertarget{wood-density-and-tree-architectural-diversity}{%
\subsection{Wood density and tree architectural
diversity}\label{wood-density-and-tree-architectural-diversity}}

@@

{\centering \includegraphics{../figs/wd_com.png}

@@

\end{figure}

@@

{\centering \includegraphics{../figs/wd_ef.png}

@@

\end{figure}

\hypertarget{estimating-aboveground-biomass}{%
\subsection{Estimating aboveground
biomass}\label{estimating-aboveground-biomass}}

@@

{\centering \includegraphics{../figs/agb_ang.png}

@@

\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\hypertarget{tree-allometry}{%
\subsection{Tree allometry}\label{tree-allometry}}

>>>>

\begin{document}

@@

\section{Introduction}\label{introduction}

Mapping terrestrial carbon stocks is crucial to successfully
implementing climate change mitigation programs
(\citeproc{ref-Chave2014}{Chave et al. 2014}). The accuracy of carbon
stock estimates depends on the availability of reliable allometric
models to deduce the aboveground biomass (AGB) of trees from census data
(\citeproc{ref-Chave2014}{Chave et al. 2014}). Selecting an appropriate
allometric model is, therefore, a critical step in estimating forest
carbon biomass (e.g., \citeproc{ref-Chave2014}{Chave et al. 2014};
\citeproc{ref-Molto2014}{Molto et al. 2014}). Tree allometry describes
the correlations between different biometric measurements of trees, such
as diameter at breast height (DBH), tree height, and tree crown size
(e.g., \citeproc{ref-Shinozaki1964a}{Shinozaki et al. 1964};
\citeproc{ref-Niklas1994}{Niklas 1994}; \citeproc{ref-West1999}{West et
al. 1999}). These relationships are widely utilized to estimate tree
architecture from DBH and to estimate forest biomass and production,
serving as fundamental components of dynamic vegetation models (e.g.,
\citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
\citeproc{ref-Falster2017}{Falster et al. 2017};
\citeproc{ref-Song2024}{Song et al. 2024}). Despite significant
progress, developing precise allometric models across regions
characterized by diverse tree species composition and varying forest
structures remains a persistent challenge. This complexity highlights
the need for improved global-scale allometric models that capture the
diversity of tree forms and ecological contexts.

Identifying the most appropriate mathematical representation of
allometric relationships is a subject of ongoing debate. While numerous
papers have used various functional forms for allometric equations
(e.g., \citeproc{ref-Thomas1996}{Thomas 1996};
\citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
\citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019};
\citeproc{ref-Song2024}{Song et al. 2024}), power-law functions remain
the most common choice for describing the allometric scaling of tree
size variables, including height and crown size with DBH (e.g.,
\citeproc{ref-Sileshi2023}{Sileshi et al. 2023};
\citeproc{ref-Jucker2024}{Jucker et al. 2024}). The relationships among
tree size variables often involve complex interactions influenced by
various factors, such as species diversity, environmental conditions,
and individual variability (\citeproc{ref-Laurans2024}{Laurans et al.
2024}; e.g., \citeproc{ref-Poorter2012}{\textbf{Poorter2012?}}).
Consequently, power-law models often fail to accurately capture the
allometric relationships across the full range of tree sizes, leading to
underestimation of dimensions for smaller trees and underestimation for
larger trees (e.g., \citeproc{ref-Fayolle2016}{Fayolle et al. 2016};
\citeproc{ref-Song2024}{Song et al. 2024}). This highlights the need to
adopt alternative functional forms---such as saturating
relationships---that can more accurately reflect the complexity of
allometric scaling in vegetation models (e.g.,
\citeproc{ref-Goodman2014}{Goodman et al. 2014};
\citeproc{ref-Song2024}{Song et al. 2024}).

Earth observation (EO) technologies, including remote sensing and Light
Detection and Ranging (LiDAR), have revolutionized monitoring forests
(e.g., \citeproc{ref-Jucker2017}{Jucker et al. 2017}) by enabling the
identification and measurement of individual tree heights and crown
dimensions using airborne imagery (e.g.,
\citeproc{ref-Shendryk2016}{Shendryk et al. 2016};
\citeproc{ref-Jucker2017}{Jucker et al. 2017}). Because existing
allometric models still depend on stem diameter as a crucial input for
biomass estimation, it also presents the challenge of determining the
most effective way to utilize EO technologies for AGB estimation
(\citeproc{ref-Shendryk2016}{Shendryk et al. 2016};
\citeproc{ref-Jucker2017}{Jucker et al. 2017}). While ground-based LiDAR
can measure DBH directly and is much more efficient than traditional
ground surveys (\citeproc{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is
still less common and convenient than aerial LiDAR. Alternative
approaches focusing on tree height and crown dimensions as model
predictors are also necessary for remote sensing applications
(\citeproc{ref-Jucker2017}{Jucker et al. 2017}). One potential solution
is to estimate DBH from tree height and crown size and then use
established biomass equations, which could be an effective strategy for
integrating airborne imagery into carbon monitoring programs
(\citeproc{ref-Jucker2017}{Jucker et al. 2017}).

Accurately representing tree allometry relationships also requires
considering underlying ecological constraints, particularly the
interplay of plant functional traits and their trade-offs influences
individual tree architecture and the composition of entire forest
communities (e.g., \citeproc{ref-Iida2012}{Iida et al. 2012};
\citeproc{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
\citeproc{ref-Yang2023}{Yang and Swenson 2023}). Under closed-canopy
conditions, strong light competition makes the balance between vertical
and horizontal space use crucial for species coexistence (e.g.,
\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\citeproc{ref-Iida2012}{Iida et al. 2012}). Among functional traits,
wood density is closely linked to both plant life history strategies
(\citeproc{ref-Fajardo2022}{\textbf{Fajardo2022?}}) and light demand
(e.g., \citeproc{ref-King2006}{King et al. 2006};
\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\citeproc{ref-Poorter2012}{\textbf{Poorter2012?}}). Several studies
suggest that wood density explains a trade-off between rapid stem
expansion and broader crowns, shaping mechanical constraints on species
with diverse architecture and light capture strategies
(\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\citeproc{ref-Iida2012}{Iida et al. 2012}; \citeproc{ref-Yang2023}{Yang
and Swenson 2023}). However, these types of studies have examined this
relationship post hoc, relating species-specific allometric parameters
to wood density and risking Type I or Type II errors if the hierarchical
data structure is overlooked (\citeproc{ref-Gelman2006}{Gelman and Hill
2006}). One exception is Martínez Cano et al.
(\citeproc{ref-MartinezCano2019}{2019}), who incorporated wood density
into a hierarchical model of tropical trees and concluded it offered
minimal explanatory power for interspecific variation. Therefore, the
role of wood density in global allometric scaling remains insufficiently
explored.

In this study, we present a robust Bayesian hierarchical approach to
delineate allometric relationships involving DBH, tree height (H), crown
radius (CR), and their interspecific variations. Using large datasets
from the Tallo and TRY databases (\citeproc{ref-Kattge2020}{Kattge et
al. 2020}; \citeproc{ref-Jucker2022}{Jucker et al. 2022}), which span a
diverse array of tree forms across global forest types and climatic
conditions, we overcome the limitations of previous studies that often
relied on site-specific or regional observations with limited samples
and species. By integrating species-level wood density into a multilevel
model, our methodology avoids the post-hoc correlations used in previous
studies, offering a more universal and scalable model, effectively
capturing variability from the species to the community level.
Specifically, we ask the following questions: (1) What is the best
function to estimate tree architecture from DBH and \emph{vice versa}?
(2) How does wood density shape interspecific variation in these
allometric scaling relationships? (3) How does the choice of allometric
relationships affect subsequent AGB estimation?

\section{Materials and Methods}\label{materials-and-methods}

\subsection{Data}\label{data}

Our analysis combined tree allometry data from Tallo, which includes
standardized measurements of DBH, height, crown radius for 5,163 species
across various biomes (\citeproc{ref-Jucker2022}{Jucker et al. 2022},
https://doi.org/10.5281/zenodo.6637599) along with wood density data for
10,995 species from TRY version 6 (\citeproc{ref-Kattge2020}{Kattge et
al. 2020}, https://www.try-db.org/TryWeb/Home.php) (see Supplementary
Table S1 for further details). We thoroughly processed the Tallo dataset
by removing 998 outliers following Jucker et al.
(\citeproc{ref-Jucker2022}{2022}). To ensure consistency, we selected
only species with standardized wood density values in g
cm\textsuperscript{-3} from TRY and removed any records with missing
data. Subsequently, we created an overlapping dataset between Tallo and
TRY, resulting in a combined dataset encompassing allometric and wood
density information for 355,881 individuals across 2,961 species.

To reduce computational complexity and ensure the robustness of the
models, we further filtered species with fewer than 20 individuals and
randomly sampled up to 100 individuals per species. This step was
necessary because the dataset is unbalanced; for example, \emph{Quercus
ilex} alone accounts for over 20,000 individuals, representing
approximately 10\% of the dataset. However, \emph{Quercus ilex} is not
necessarily the most abundant species globally. Its prevalence in this
dataset is a result of sampling and does not reflect its true global
abundance. By limiting the number of individuals per species, we
minimize bias due to data imbalance and prevent the overrepresentation
of any single abundant species (\citeproc{ref-Gelman2006}{Gelman and
Hill 2006}), thereby ensuring that the model remains broadly applicable.
Since not all trees have a complete set of information, we created
specific sub-datasets to select the best predictive model for each
allometric relationship for two clades (i.e., angiosperms and
gymnosperms; see below and Supplementary Table S2 for more details).

\subsection{Tree allometry models}\label{tree-allometry-models}

@@

To compare the performance of different models for the same response
variable, we used Pareto-smoothed importance sampling leave-one-out
cross-validation (PSIS-LOO; Vehtari et al.
(\citeproc{ref-Vehtari2017}{2017})). PSIS-LOO is an accurate and
reliable approximation to standard leave-one-out cross-validation (LOO),
which is a robust method for comparing models with different numbers of
parameters (\citeproc{ref-Vehtari2017}{Vehtari et al. 2017}). We
utilized PSIS-LOO to calculate the LOO Information Criterion (LOOIC) for
each model. A model with lower LOOIC is considered to have better
predictive accuracy (\citeproc{ref-Vehtari2017}{Vehtari et al. 2017}).

Once the best-performing models were identified for each allometric
relationship, we then tested whether adding a species-level predictor
(wood density) would further improve that model's predictive
performance. Species-specific parameters (\(\beta_j\)) were modeled as a
function of wood density (\(u_j\)):
\begin{equation}\phantomsection\label{eq-wood}{
\beta_{j} \sim N(\gamma_{0} + \gamma_{1} u_j, \sigma_{\beta}),
}\end{equation}

@@

Posterior distributions of all parameters were estimated using the
No-U-Turn Sampler (NUTS), an adaptive variant of the Hamiltonian Monte
Carlo (HMC), algorithm implemented in Stan
(\citeproc{ref-Carpenter2017}{Carpenter et al. 2017}) using
weakly-informative priors (\citeproc{ref-Gelman2008}{Gelman et al.
2008}). The HMC algorithm uses gradient information to propose new
states in the Markov chain, leading to a more efficient exploration of
the target distribution than traditional Markov chain Monte Carlo (MCMC)
methods that rely on random proposals
(\citeproc{ref-Carpenter2017}{Carpenter et al. 2017}). This efficiency
allows us to achieve convergence with fewer iterations than traditional
MCMC methods. Inferences were based on 1,000 posterior samples following
1,000 warmup iterations for three parallel chains. Convergence was
assessed using the Gelman--Rubin statistic with a convergence threshold
of 1.1 (\citeproc{ref-Gelman2013}{Gelman et al. 2015}) and ensuring
effective sample sizes greater than 300
(\citeproc{ref-Vehtari2021}{Vehtari et al. 2021}). We also reported the
median and 95\% Bayesian credible intervals (CIs) for each parameter
(Supplementary Tables S3-S8). Estimates were considered significant when
95\% CIs did not overlap with zero.

Statistical analyses were performed in R version 4.3.2
(\citeproc{ref-RCoreTeam2023}{R Core Team 2023}), using the R packages
\emph{targets} version 1.7.0 for workflow management
(\citeproc{ref-Landau2021}{Landau 2021}). Codes are made available on
GitHub at \url{https://github.com/duyenecology/tree-allometry} (Stan
code is provided in Supplementary Section S2).

\subsection{Aboveground biomass
estimates}\label{aboveground-biomass-estimates}

For angiosperm species, we estimated aboveground biomass (AGB; kg of dry
mass) using the equation from Chave et al.
(\citeproc{ref-Chave2014}{2014}):

\begin{equation}\phantomsection\label{eq-agb}{
\mathrm{AGB} = 0.0673 \times (\rho D^2H)^{0.976}
}\end{equation}

where \(\rho\) is wood density (g cm\textsuperscript{-3}), \(D\) (cm) is
DBH, and \(H\) (m) is tree height. We either used both observed DBH and
height or substituted one of these variables with an estimate from our
allometric models, allowing us to assess how different allometric
relationships affect AGB calculations. Only trees with a DBH ≥ 5 cm were
used, as those below this threshold hold a small fraction of forest
carbon stocks and were excluded from the calibration of Chave et al.
(\citeproc{ref-Chave2014}{2014})'s equation. Additionally, the
sub-datasets used to select the optimal predictive model for each
allometric relationship differed as not all trees had a complete
information set (see above and Supplementary Table S2). Consequently,
the final dataset for estimating AGB in angiosperms includes 124,206
trees with sufficient allometric information (i.e., DBH, H, CR, WD)
across 624 species. We did not extend this approach to gymnosperms
because no general AGB equation is available for them, and most existing
equations are species-specific and do not incorporate wood density
(e.g., \citeproc{ref-Moore2010}{Moore 2010};
\citeproc{ref-Bazrgar2024}{Bazrgar et al. 2024}).

\section{Results}\label{results}

\subsection{Tree height allometry}\label{tree-height-allometry}

The optimal model to predict tree height was a Weibull function for both
angiosperms (ang) and gymnosperms (gym) (Table 1). Across the global
dataset, the best model to predict H from DBH of each division was:
\begin{equation}\phantomsection\label{eq-h_ang}{
\mathrm{H_{ang}} =50.0
\left\{ 1 - \exp\left( - 0.0476
\mathrm{DBH}^{0.629}
\right) \right\}
}\end{equation}

\begin{equation}\phantomsection\label{eq-h_gym}{
\mathrm{H_{gym}} = 117.4
\left\{ 1 - \exp\left( - 0.012
\mathrm{DBH}^{0.750}
\right) \right\}.
}\end{equation}

@@

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../figs/h_cr_dbh_non_log.png}}

@@

\end{figure}%

\subsection{Crown Radius allometry}\label{crown-radius-allometry}

The power-law function (0 \textless{} \emph{b} \textless{} 1) provided
the best predictive performance for modeling the relationship between
crown radius and DBH in angiosperms. Whereas for gymnosperms, the gMM
function offered the best performance for this allometric scaling (Table
1). Across the global dataset, the best model for predicting CR from DBH
of each division was: \begin{equation}\phantomsection\label{eq-cr_ang}{
\mathrm{CR_{ang}} = \mathrm{0.349}
\, \mathrm{DBH}^{\mathrm{0.626}}
}\end{equation}

\begin{equation}\phantomsection\label{eq-cr_gym}{
\mathrm{CR_{gym}} = \frac{\mathrm{769.4}
\, \mathrm{DBH}^{\mathrm{0.631}}}
{\mathrm{2657}
+ \mathrm{DBH}^{\mathrm{0.631}}}.
}\end{equation}

@@

\subsection{DBH allometry}\label{dbh-allometry}

The \(CR,H\) model, which treated CR and H as separate factors,
exhibited the best predictive performance among all tested models (Table
1). Across the global dataset, the most effective model for predicting
DBH from H and CR for each division was:
\begin{equation}\phantomsection\label{eq-dbh_ang}{
\mathrm{DBH_{ang}} = \mathrm{0.948}
\mathrm{CR}^{\text{0.374}}
\mathrm{H}^{\text{0.590}}
}\end{equation}

\begin{equation}\phantomsection\label{eq-dbh_gym}{
\mathrm{DBH_{gym}} = \mathrm{0.972}
\mathrm{CR}^{\text{0.410}}
\mathrm{H}^{\text{0.588}}.
}\end{equation}

@@

\subsection{Wood density and tree architectural
diversity}\label{wood-density-and-tree-architectural-diversity}

@@

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../figs/wd_com.png}}

@@

\end{figure}%

@@

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../figs/wd_ef.png}}

@@

\end{figure}%

\subsection{Estimating aboveground
biomass}\label{estimating-aboveground-biomass}

@@

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../figs/agb_ang.png}}

@@

\end{figure}%

\section{Discussion}\label{discussion}

@@

\subsection{Tree allometry}\label{tree-allometry}
